A Survey of Quantitative Descriptions of Molecular Structure

### 2012-04-08 paper.tex
Added some more text on shape descriptors

A number of geometric descriptors characterize the molecular
surface. The simplest such example is the surface area descriptor (and
the corresponding volume descriptor). A simple extension of such a
descriptor is to characterize distributions of a physicochemical
property over the molecular surface such as the Charged Partial
Surface Area descriptors \cite{Stanton:1990aa} that characterize the
partial charge distribution over the molecular surfce. Note that
surface descriptor calculations based on analytical
\cite{Connolly:1983aa} or tesselation algorithms
\cite{Eisenhaber:1995qf} can be slow and that parametrized methods
such as the VSA descriptors \cite{Labute:2008aa} that are based on
precalculated surface area values derived from a list of functional
groups. A comparison of surface area values generated by the two
approaches for XXX molecules indicates a high degree of correlation.
\textbf{Ray tracing descriptors, ultrafast shape descriptors,
  spherical harmonic shape descriptors, field based descriptors
  [Cresset etc]}

>>>>

A number of geometric descriptors characterize the molecular
surface. The simplest such example is the surface area descriptor (and
the corresponding volume descriptor). A simple extension of such a
descriptor is to characterize distributions of a physicochemical
property over the molecular surface such as the Charged Partial
Surface Area descriptors \cite{Stanton:1990aa} that characterize the
partial charge distribution over the molecular surfce. Note that
surface descriptor calculations based on analytical
\cite{Connolly:1983aa} or tesselation algorithms
\cite{Eisenhaber:1995qf} can be slow and that parametrized methods
such as the VSA descriptors \cite{Labute:2008aa} that are based on
precalculated surface area values derived from a list of functional
groups. A comparison of surface area values generated by the two
approaches for XXX molecules indicates a high degree of
correlation.

While molecular surfaces are useful, molecular shape is a key factor
in ligand-receptor interactions. A variety of descriptors have been
devised to characterize molecular shape. For example, the ``shape
signature'' approach described by Zauhar et al \cite{Zauhar:2003fk} is a novel application of a
ray tracing method that generates a description of the molecular
shape. Ballester et al \cite{Ballester:2007aa} described the Ultrafast
Shape Recognition (USR) approach that characterized molecular shape in
terms of the distribution of distances between the atoms and a set of
pre-defined points. As a result, the calculation procedure is
extremely fast and once calculated shape similarities can be evaluated
rapidly via a simple distance calculation. 
\textbf{
  spherical harmonic shape descriptors, field based descriptors
  [Cresset etc]}
